---
output:
  pdf_document: default
  html_document: default
---
# EDA

## Set up the enviroment

```{r}
url <- "https://github.com/DavidVarga1995/understat_data/archive/refs/heads/main.zip"
dest <- "./understat_data.zip"

download.file(url, dest)
unzip(dest)
```

```{r}
library(dplyr)
```

Read the data

```{r}
epl_shots <- read.csv('./understat_data-main/data/Shots/EPL_shots_2020.csv')
epl_players <- read.csv('./understat_data-main/data/Players/EPL_players_2020.csv')
epl_games <- read.csv('./understat_data-main/data/Games/EPL_games_2020.csv')
epl_fixtures <- read.csv('./understat_data-main/data/Fixtures/EPL_fixtures_2020.csv')
bundesliga_shots <- read.csv('./understat_data-main/data/Bundesliga_2019/Bundesliga_shots_2019.csv')
```

Descriptive Statistics

```{r}
bundesliga_shots_attributes <- select(bundesliga_shots, 'minute', 'h_goals', 'a_goals', 'situation', 'lastAction', 'X', 'Y', 'shotType', 'xG', 'result')

print(summary(bundesliga_shots_attributes))
```

Summarize factors

```{r}
table(bundesliga_shots_attributes$situation)
```

As we can see the most common situation of a shot is open play. The second most common situation of a shot is corner.

The frequency of set pieces and direct freekicks is significant, too.

However the frequency of penalties as a situation of a shot is insignificant.

```{r}
table(bundesliga_shots_attributes$lastAction)
```

Obviously the most common action before a shot is a pass. However the frequency of crosses and aerials is significant, too.

The second most common action before a shot is none - resp. the shot is immediately after a ball recovery.

The frequency of standard situations (set pieces), chipped balls, take-ons, rebounds, ball recoveries, head passes, through balls and ball touches is also notable.

The frequency of the remaining values is insignificant.

```{r}
table(bundesliga_shots_attributes$shotType)
```

As expected the majority of shots is shot by right or left foot. However the frequency of headers is significant, too.

The remaning shots are shot with other body part.

```{r}
table(bundesliga_shots_attributes$result)
```

The result of the shot most frequently is a miss or a blocked shot. However the the amount of saved shots is almost equal with the amount of saved shots.

If a shot is not missed, blocked or saved by the goalkeeper it can be a goal or it can hit the post.

The amount of own goals is insignificant.

```{r}
table(bundesliga_shots_attributes$result, bundesliga_shots_attributes$shotType)
```

```{r}
table(bundesliga_shots_attributes$lastAction, bundesliga_shots_attributes$shotType)
```

```{r}
table(bundesliga_shots_attributes$situation, bundesliga_shots_attributes$shotType)
```

```{r}
table(bundesliga_shots_attributes$lastAction, bundesliga_shots_attributes$result)
```

```{r}
table(bundesliga_shots_attributes$situation, bundesliga_shots_attributes$result)
```

```{r}
table(bundesliga_shots_attributes$lastAction, bundesliga_shots_attributes$situation)
```

```{r}
print(aggregate(xG ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(xG ~ situation, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(xG ~ lastAction, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(xG ~ result, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(minute ~ lastAction, data = bundesliga_shots_attributes, summary))
print(aggregate(minute ~ result, data = bundesliga_shots_attributes, summary))
print(aggregate(minute ~ situation, data = bundesliga_shots_attributes, summary))
print(aggregate(minute ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(X ~ lastAction, data = bundesliga_shots_attributes, summary))
print(aggregate(X ~ result, data = bundesliga_shots_attributes, summary))
print(aggregate(X ~ situation, data = bundesliga_shots_attributes, summary))
print(aggregate(X ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(Y ~ lastAction, data = bundesliga_shots_attributes, summary))
print(aggregate(Y ~ result, data = bundesliga_shots_attributes, summary))
print(aggregate(Y ~ situation, data = bundesliga_shots_attributes, summary))
print(aggregate(Y ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(h_goals ~ lastAction, data = bundesliga_shots_attributes, summary))
print(aggregate(h_goals ~ result, data = bundesliga_shots_attributes, summary))
print(aggregate(h_goals ~ situation, data = bundesliga_shots_attributes, summary))
print(aggregate(h_goals ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
print(aggregate(a_goals ~ lastAction, data = bundesliga_shots_attributes, summary))
print(aggregate(a_goals ~ result, data = bundesliga_shots_attributes, summary))
print(aggregate(a_goals ~ situation, data = bundesliga_shots_attributes, summary))
print(aggregate(a_goals ~ shotType, data = bundesliga_shots_attributes, summary))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$minute), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$minute), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$minute), "\n")
cat("Range: ", range(bundesliga_shots_attributes$minute), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$minute))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$xG), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$xG), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$xG), "\n")
cat("Range: ", range(bundesliga_shots_attributes$xG), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$xG))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$X), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$X), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$X), "\n")
cat("Range: ", range(bundesliga_shots_attributes$X), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$X))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$Y), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$Y), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$Y), "\n")
cat("Range: ", range(bundesliga_shots_attributes$Y), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$Y))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$h_goals), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$h_goals), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$h_goals), "\n")
cat("Range: ", range(bundesliga_shots_attributes$h_goals), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$h_goals))
```

```{r}
cat("Variance: ", var(bundesliga_shots_attributes$a_goals), "\n")
cat("Standard Deviation: ", sd(bundesliga_shots_attributes$a_goals), "\n")
cat("Quantiles: ", quantile(bundesliga_shots_attributes$a_goals), "\n")
cat("Range: ", range(bundesliga_shots_attributes$a_goals), "\n")
cat("Interquartile Range: ", IQR(bundesliga_shots_attributes$a_goals))
```

```{r}
bundesliga_shots_attributes$situation <- as.factor(bundesliga_shots_attributes$situation)
bundesliga_shots_attributes$shotType <- as.factor(bundesliga_shots_attributes$shotType)
bundesliga_shots_attributes$result <- as.factor(bundesliga_shots_attributes$result)
bundesliga_shots_attributes$lastAction <- as.factor(bundesliga_shots_attributes$lastAction)
```

```{r}
str(bundesliga_shots_attributes)
```

```{r}
vars <- c('minute', 'h_goals', 'a_goals', 'X', 'Y', 'xG')

round(cor(bundesliga_shots_attributes[, vars], use = "complete.obs"), 2)
```

```{r}
library(polycor)
```

```{r}
hetcor(bundesliga_shots_attributes)
```

```{r}
library(ggplot2)

ggplot(bundesliga_shots_attributes, aes(x = h_goals)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = a_goals)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = situation)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = shotType)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = result)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = minute)) +
  geom_histogram(bins = nclass.Sturges(bundesliga_shots_attributes$minute))
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X)) +
  geom_histogram(bins = nclass.Sturges(bundesliga_shots_attributes$X))
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y)) +
  geom_histogram(bins = nclass.Sturges(bundesliga_shots_attributes$Y))
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = xG)) +
  geom_histogram(bins = nclass.Sturges(bundesliga_shots_attributes$xG))
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = minute)) +
  geom_density()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X)) +
  geom_density()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y)) +
  geom_density()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = xG)) +
  geom_density()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = situation, y = minute))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = shotType, y = minute))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = result, y = minute))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = situation, y = xG))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = shotType, y = xG))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = result, y = xG))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = situation, y = X))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = shotType, y = X))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = result, y = X))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = situation, y = Y))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = shotType, y = Y))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, (aes(x = result, y = Y))) +
  geom_boxplot()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(sample = xG)) + 
  geom_qq() +
  geom_qq_line()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(sample = X)) + 
  geom_qq() +
  geom_qq_line()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(sample = Y)) + 
  geom_qq() +
  geom_qq_line()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(sample = minute)) + 
  geom_qq() +
  geom_qq_line()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = X)) +
  geom_point()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = xG)) +
  geom_point()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, y = xG)) +
  geom_point()
```

```{r}
library(GGally)
```

```{r}
vars <- c('minute', 'X', 'Y', 'xG')

ggpairs(bundesliga_shots_attributes[vars])
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = X, color = result)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = X, color = shotType)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = X, color = situation)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = xG, color = result)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = xG, color = shotType)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, y = xG, color = situation)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, y = xG, color = result)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, y = xG, color = situation)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, y = xG, color = shotType)) +
  geom_point(size = 3)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = xG, color = result)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = xG, color = shotType)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, color = situation)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, color = result)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = Y, color = shotType)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, color = result)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, color = situation)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = X, color = shotType)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = minute, color = result)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = minute, color = shotType)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = minute, color = situation)) +
  geom_density(size = 2)
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = shotType, fill = result)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = situation, fill = result)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = shotType, fill = situation)) +
  geom_bar()
```

```{r}
ggplot(bundesliga_shots_attributes, aes(x = result, y = situation)) + 
  geom_tile(alpha = 0.01) 
```

```{r}

```

